---
title: "Business Indicators"
format:
  html:
    theme: cosmo
---

## 

Please [contact us](mailto:regional@upjohn.org) if you have any questions. 

## 


```{r}
#| label: Personal Consumption
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

spending <- readRDS("data/fred_spending.rds")

fig <- plot_ly(spending, x = ~date)

fig <- fig %>% 
  add_lines(
    y = ~disposable_income, 
    name = "Disposable Income",
    line =list(color = "#004A5E"),
    hovertemplate = "Disposable Income: %{y:,}<extra></extra>"
  )
fig <- fig %>% 
  add_lines(y = ~durable_goods, 
            name = "Durable Goods",
            line = list(color = "#2B9CB4"),
            hovertemplate = "Durable Goods: %{y:,}<extra></extra>"
  )
fig <- fig %>% 
  add_lines(y = ~nondurable_goods, 
            name = "Nondurable Goods",
            line = list(color = "#AF2C32"),
             hovertemplate = "Nondurable Goods: %{y:,}<extra></extra>"
  )
fig <- fig %>% 
  add_lines(y = ~services, 
            name = "Services",
            line = list(color = "#414042"),
            hovertemplate = "Services: %{y:,}<extra></extra>"
            )

fig <- fig %>%
  layout(
    title = list(
      text = paste(
        "Real Personal Consumption Expenditures and Disposable Personal Income",
        "<br><sup>Seasonally Adjusted</sup>"
      )
    ),
    shapes = list(
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2007-12-01",
        x1 = "2009-06-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      ),
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2020-02-01",
        x1 = "2020-04-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      )
    ),
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Billions of Chained 2017 Dollars"
    ),
    annotations = list(
      list(
        x = "2008-09-01",
        y = 0.95,
        xref = "x",
        yref = "paper",
        text = "Great Recession",
        showarrow = FALSE,
        font = list(size = 10)
      ),
    
      list(
        x = "2020-03-01",
        y = 0.95,
        xref = "x",
        yref = "paper",
        text = "COVID Recession",
        showarrow = FALSE,
        font = list(size = 10)
      )
    ),
    margin = list(
      t = 80
    )
  )

config(
  fig,
  displayModeBar = FALSE
)
```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Real personal consumption expenditures and disposable income are crucial for understanding consumer spending patterns, which drive much of economic activity. These indicators reveal household financial health and confidence, guiding policymakers in assessing fiscal policies and fostering economic growth and stability. "Chained 2017 dollars" is a method of adjusting economic statistics for inflation; values from other years are adjusted to reflect purchasing power of the dollar in 2017. Consumer goods can be classified as durable, nondurable, or services. Durable goods are classified as items that have long periods between successive purchases (ex. cars or home appliances). Nondurable goods are classified as items that have long periods between successive purchases (ex. food or clothing). Services are acts which are paid for (ex. work done by barbers, doctors, mechanics, etc.) 

**Source:** Federal Reserve Bank of St. Louis (series DSPIC96, PCEDGC96, PCENDC96, and PCESC96)

[Download Data](data_download/fred_spending.csv){download="fred_spending.csv"}
:::

##

::: {.plot-space}
:::

```{r}
#| label: Vehicle Sales and Fuel Prices
#| echo: false
#| message: false
#| warning: false

library(plotly)

autoFuel <- readRDS("data/vehicleSalesFuelPrices.rds") %>%
  filter(date >= as.Date("2007-01-01")) %>%
  mutate(
    value.y = as.numeric(value.y)
  )

fig <- plot_ly(autoFuel, x = ~date)

fig <- fig %>%
  add_lines(
    y = ~value.x,
    name = "Light Vehicle Sales",
    yaxis = "y1",
    line =list(color = "#004A5E")
  )

fig <- fig %>%
  add_lines(
    y = ~value.y,
    name = "Fuel Prices",
    yaxis = "y2",
    line =list(color = "#AF2C32")
  )

fig %>%
  layout(
    title = list(
      text = "Light Vehicle Sales and Fuel Prices<br><sup>United States, Monthly Data Since 2007</sup>"
    ),
    
    shapes = list(
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2007-12-01",
        x1 = "2009-06-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      ),
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2020-02-01",
        x1 = "2020-04-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      )
    ),

    xaxis = list(
      title = "Date",
      showline = TRUE,
      linewidth = 1
    ),

    yaxis = list(
      title = "Light Vehicle Sales (Millions)",
      side = "left",
      range = c(1,20)
    ),

    yaxis2 = list(
      title = "Fuel Prices ($ per Gallon)",
      overlaying = "y",
      side = "right",
      range = c(1,20)
    ),

    legend = list(
      x = 1.05,
      y = 1,
      xanchor = "left"
    ),

    annotations = list(
       list(
        x = "2008-09-01",
        y = 0.95,
        xref = "x",
        yref = "paper",
        text = "Great Recession",
        showarrow = FALSE,
        font = list(size = 10)
      ),
    
      list(
        x = "2020-03-01",
        y = 0.95,
        xref = "x",
        yref = "paper",
        text = "COVID Recession",
        showarrow = FALSE,
        font = list(size = 10)
      )
    ),

    margin = list(
      t = 60,
      r = 120
    )
  ) %>%
  config(displayModeBar = FALSE)


```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Vehicle sales and fuel prices are important economic indicators because they reflect consumer spending and economic confidence, with high sales signaling robust economic activity. Additionally, fuel prices impact transportation costs and disposable income, influencing the economy overall.

**Source:**  Federal Reserve Bank of St. Louis (series ALTSALES) and U.S. Energy Information Administration (series EMM_EPM0_PTE_NUS_DPG)

[Download Data](data_download/vehicleSalesFuelPrices.csv){download="vehicleSalesFuelPrices.csv"}
:::

##

::: {.plot-space}
:::

```{r}
#| label: GDP Percent Change
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)

# ---------------------------------------------------------
# Read and prepare data
# ---------------------------------------------------------

gdp <- readRDS("data/gdpPercentChange.rds") %>%
  mutate(
    TimePeriod = as.Date(
      paste0(TimePeriod, "-01-01")
    )
  )

# Michigan data
mi_gdp <- gdp %>%
  select(
    TimePeriod,
    Michigan_GDP_percent_change
  ) %>%
  distinct() %>%
  arrange(TimePeriod)

# Counties
counties <- unique(gdp$GeoName)

init <- counties[1]

# ---------------------------------------------------------
# Create plot
# ---------------------------------------------------------

p <- plot_ly()

# Add one bar trace for each county
for (cty in counties) {

  df <- gdp %>%
    filter(GeoName == cty) %>%
    arrange(TimePeriod)

  p <- p %>%
    add_bars(
      data = df,
      x = ~TimePeriod,
      y = ~GDP_percent_change,

      hovertemplate = paste0(
        "County: ", cty,
        "<br>Year: %{x|%Y}",
        "<br>GDP Change: %{y:.2f}%",
        "<extra></extra>"
      ),

      name = cty,

      visible = (cty == init),

      marker = list(
        color = "#004A5E"
      ),

      legendgroup = "county"
    )
}

# ---------------------------------------------------------
# Add Michigan bars
# ---------------------------------------------------------

p <- p %>%
  add_bars(
    data = mi_gdp,
    x = ~TimePeriod,
    y = ~Michigan_GDP_percent_change,

    hovertemplate = paste0(
      "Michigan",
      "<br>Year: %{x|%Y}",
      "<br>GDP Change: %{y:.2f}%",
      "<extra></extra>"
    ),

    name = "Michigan",

    marker = list(
      color = "#2B9CB4"
    ),

    inherit = FALSE
  )

# ---------------------------------------------------------
# County dropdown
# ---------------------------------------------------------

buttons <- lapply(seq_along(counties), function(i) {

  # Hide every county
  vis <- rep(
    FALSE,
    length(counties) + 1
  )

  # Show selected county
  vis[i] <- TRUE

  # Michigan is always visible
  vis[length(vis)] <- TRUE

  list(
    method = "update",

    args = list(
      list(
        visible = vis
      ),

      list(
        title = paste(
          "GDP Percent Change:",
          counties[i],
          "vs Michigan"
        )
      )
    ),

    label = counties[i]
  )
})

# ---------------------------------------------------------
# Layout
# ---------------------------------------------------------

p %>%
  layout(

    title = paste(
      "GDP Percent Change:",
      init,
      "vs Michigan"
    ),

    # Place county and Michigan bars side-by-side
    barmode = "group",

    xaxis = list(
      title = "Year",
      type = "date",
      tickformat = "%Y",
      dtick = "M12"
    ),

    yaxis = list(
      title = "GDP Percent Change",
      ticksuffix = "%"
    ),

    updatemenus = list(
      list(
        type = "dropdown",
        buttons = buttons,
        x = 1.4,
        y = 0.75,
        xanchor = "right"
      )
    ),

    legend = list(
      x = 1.05,
      y = 1,
      xanchor = "left"
    ),

    margin = list(
      r = 200,
      t = 30
    )

  ) %>%
  config(
    displayModeBar = FALSE
  )

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

The GDP percent change measures the economy's growth rate, reflecting overall economic health and stability. It helps policymakers and investors make informed decisions by indicating the pace of economic expansion or contraction.

**Source:** US BEA (table CAGDP1)

[Download Data](data_download/gdpPercentChange.csv){download="gdpPercentChange.csv"}
:::

##

::: {.plot-space}
:::

```{r}
#| label: Business Applications Change From a Year Ago
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)

biz <- readRDS("data/businessAppChange.rds") %>%
  mutate(Date = as.Date(date)) %>%
  filter(format(Date, "%Y") != "2006") %>%
  arrange(Date)

biz <- biz %>%
  mutate(
    Date = as.Date(date),
    Year = format(Date, "%Y")
  )

# Get years
years <- sort(unique(biz$Year), decreasing = TRUE)

# Full date range
all_date_min <- min(biz$Date, na.rm = TRUE)
all_date_max <- max(biz$Date, na.rm = TRUE)

# All Years button
all_button <- list(
  method = "update",
  
  args = list(
    
    # Restore all data
    list(
      x = list(
        biz$Date,
        biz$Date
      ),
      y = list(
        biz$Michigan,
        biz$United_States
      )
    ),
    
    # Restore full chart range
    list(
      title = "Percent Change in Business Applications",
      xaxis = list(
        range = c(
          all_date_min,
          all_date_max
        )
      )
    )
  ),
  
  label = "All Years"
)

# Year buttons
year_buttons <- lapply(years, function(selected_year) {
  
  filtered_data <- biz %>%
    filter(Year == selected_year)
  
  date_min <- min(
    filtered_data$Date,
    na.rm = TRUE
  )
  
  date_max <- max(
    filtered_data$Date,
    na.rm = TRUE
  )
  
  list(
    method = "update",
    
    args = list(
      
      # Update data
      list(
        x = list(
          filtered_data$Date,
          filtered_data$Date
        ),
        y = list(
          filtered_data$Michigan,
          filtered_data$United_States
        )
      ),
      
      # Update title and x-axis
      list(
        title = paste(
          "Percent Change in Business Applications:",
          selected_year
        ),
        xaxis = list(
          range = c(
            date_min,
            date_max
          )
        )
      )
    ),
    
    label = selected_year
  )
})

# Combine All Years + individual years
buttons <- c(
  list(all_button),
  year_buttons
)

fig <- plot_ly(biz, x = ~Date)

fig <- fig %>%
  add_lines(
    y = ~Michigan,
    name = "Michigan",
    line = list(
      color = "#004A5E"
    ),
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "Michigan: %{y:.2f}%",
      "<extra></extra>"
    )
  )

fig <- fig %>%
  add_lines(
    y = ~United_States,
    name = "United States",
    line = list(
      color = "#AF2C32",
      dash = "dash"
    ),
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "United States: %{y:.2f}%",
      "<extra></extra>"
    )
  )

fig %>%
  layout(
    title = list(
      text = "Percent Change in Business Applications<br><sup>Michigan vs United States</sup>"
    ),

    xaxis = list(
      title = "",
      showline = TRUE
    ),

    yaxis = list(
      title = "Percent Change from Year Ago",
      zeroline = TRUE,
      zerolinecolor = "gray"
    ),

    legend = list(
      x = 1.05,
      y = 1,
      xanchor = "left"
    ),
    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.05,
        y = 0.7,
        xanchor = "left"
      )
    ),

    shapes = list(

      # Great Recession
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2007-12-01",
        x1 = "2009-06-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      ),

      # COVID recession
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2020-02-01",
        x1 = "2020-04-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      )
    ),

    margin = list(
      t = 60
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

The percent change in business applications reflects entrepreneurial activity and economic dynamism. A rise in applications suggests increased business formation and economic growth potential, while a decline may indicate reduced economic confidence and future investment.

**Source:** Federal Reserve Bank of St. Louis (series BUSAPPWNSAMIYY and BUSAPPWNSAUS)

[Download Data](data_download/businessAppChange.csv){download="businessAppChange.csv"}
:::

##

::: {.plot-space}
:::

```{r}
#| label: Average Annual Wage
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)

# Read data
pay <- readRDS("data/averageAnnualWage.rds") %>%
  mutate(
    year_MI = as.Date(paste0(year_MI, "-01-01")),
    value_MI = as.numeric(value_MI),
    value_US = as.numeric(value_US),
    industry = sub("^NAICS\\s+", "", seriesID)
  ) %>%
  arrange(industry, year_MI)

# Industries in the order they appear in the data
industries <- unique(pay$industry)

# Initial industry
init <- industries[1]

df_init <- pay %>%
  filter(industry == init)

# Create plot
fig <- plot_ly()

# Michigan trace
fig <- fig %>%
  add_lines(
    data = df_init,
    x = ~year_MI,
    y = ~value_MI,
    name = "Michigan",
    line = list(color = "#004A5E")
  )

# United States trace
fig <- fig %>%
  add_lines(
    data = df_init,
    x = ~year_MI,
    y = ~value_US,
    name = "United States",
    line = list(color = "#2B9CB4",
                dash = "dash")
  )

# Dropdown buttons
buttons <- lapply(seq_along(industries), function(i){

  df <- pay %>%
    filter(industry == industries[i])

  list(
    method = "restyle",

    args = list(

      list(

        x = list(
          df$year_MI,
          df$year_MI
        ),

        y = list(
          df$value_MI,
          df$value_US
        )

      ),

      c(0,1)

    ),

    label = industries[i]

  )

})

# Layout
fig %>%
  layout(

    title = list(
      text = paste(
        "Average Annual Pay:",
        init
      ),
      x = 0.5,
      y = 0.88,
      xanchor = "center",
      yanchor = "top"
    ),

    xaxis = list(
      title = "",
      type = "date",
      tickformat = "%Y"
    ),

    yaxis = list(
      title = "Average Annual Pay",
      tickprefix = "$",
      tickformat = ",d",
      separatethousands = TRUE
    ),

    updatemenus = list(
      list(
        type = "dropdown",
        direction = "down",
        active = 0,
        buttons = buttons,
        x = 0.5,
        y = 1.25,
        xanchor = "center",
        yanchor = "top"
      )
    ),

    legend = list(
      orientation = "h",
      x = 0.5,
      y = 1.08,
      xanchor = "center"
    ),
    
    shapes = list(
      list(
        type = "rect",
        xref = "x",
        yref = "paper",
        x0 = "2020-02-01",
        x1 = "2020-04-01",
        y0 = 0,
        y1 = 1,
        fillcolor = "lightgray",
        opacity = 0.3,
        line = list(width = 0),
        layer = "below"
      )
    ),
    annotations = list(
    
      list(
        x = "2020-03-01",
        y = 0.95,
        xref = "x",
        yref = "paper",
        text = "COVID Recession",
        showarrow = FALSE,
        font = list(size = 10)
      )
    ),

    margin = list(
      t = 150,
      r = 40
    )

  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Tracking the average annual pay provides insights into income levels and purchasing power within an economy. It helps assess wage trends, labor market conditions, and overall economic well-being, guiding policymaking and business decisions.

**Source:** BLS QCEW

[Download Data](data_download/averageAnnualWage.csv){download="averageAnnualWage.csv"}
:::

##

::: {.plot-space}
:::

```{r}
#| label: Per Capita Income
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(knitr)
library(htmltools)
library(purrr)
library(kableExtra)



income <- readRDS(
  "data/PerCapitaIncome.rds"
)

# Southwest Michigan county order
county_order <- c(
  "Kalamazoo",
  "Berrien",
  "Branch",
  "Calhoun",
  "Cass",
  "St. Joseph",
  "Van Buren"
)

# ---------------------------------------------------------
# Create sparkline function
# ---------------------------------------------------------

make_sparkline <- function(values) {

  values <- as.numeric(values)

  # Remove missing values
  valid <- !is.na(values)

  if (sum(valid) < 2) {
    return("")
  }

  values <- values[valid]

  # Dimensions
  width <- 120
  height <- 35
  padding <- 3

  # Scale values to chart dimensions
  x <- seq(
    padding,
    width - padding,
    length.out = length(values)
  )

  value_range <- max(values) - min(values)

  if (value_range == 0) {
    y <- rep(
      height / 2,
      length(values)
    )
  } else {
    y <- height - padding -
      ((values - min(values)) / value_range) *
      (height - 2 * padding)
  }

  # Create SVG line
  points <- paste(
    paste(x, y, sep = ","),
    collapse = " "
  )

  # Return HTML SVG
  HTML(
    paste0(
      '<svg width="120" height="35" ',
      'viewBox="0 0 120 35" ',
      'xmlns="http://www.w3.org/2000/svg">',
      '<polyline ',
      'points="', points, '" ',
      'fill="none" ',
      'stroke="#2B9CB4" ',
      'stroke-width="2" ',
      'stroke-linecap="round" ',
      'stroke-linejoin="round"/>',
      '</svg>'
    )
  )
}

# ---------------------------------------------------------
# Create table data
# ---------------------------------------------------------

income_table <- income %>%
  mutate(
    
    # County ordering
    county_order = case_when(
      GeoName %in% county_order ~
        match(GeoName, county_order),
      TRUE ~
        length(county_order) + 1
    )
  ) %>%
  arrange(
    county_order,
    GeoName
  ) %>%
  mutate(
    
    # Formatted values
    `2024 Rank` = rank_2024,
    
    `Growth Rank` = pc_rank_2024,
    
    `2024 vs. 2023` = paste0(
      sprintf(
        "%.1f",
        pct_24_23
      ),
      "%"
    ),
    
Trend = pmap(
  list(
    DataValue_2020,
    DataValue_2021,
    DataValue_2022,
    DataValue_2023,
    DataValue_2024
  ),
  ~ make_sparkline(c(...))
)
  ) %>%
  select(
    GeoName,
    `2024 Rank`,
    `Growth Rank`,
    `2024 vs. 2023`,
    Trend
  )

# ---------------------------------------------------------
# Render table
# ---------------------------------------------------------

htmltools::HTML(
  '<h3 style="margin-bottom: 15px;">Per Capita Income by Michigan County</h3>'
)

htmltools::HTML(
  '<div style="margin-bottom: 15px;">',
  '<label for="countySearch"><strong>Search County:</strong></label>',
  '<input type="text" id="countySearch" ',
  'placeholder="Enter county name..." ',
  'style="margin-left: 10px; padding: 6px; width: 250px;">',
  '</div>'
)

kable(
  income_table,
  format = "html",
  escape = FALSE,
  table.attr = 'id="incomeTable"',
  col.names = c(
    "County",
    "2024 Rank",
    "Growth Rank in State",
    "2024 vs. 2023<br>Percent Change",
    "Per Capita Income:<br>2020–2024"
  ),
  align = c(
    "l",
    "c",
    "c",
    "c",
    "c"
  )
) %>%
  kable_styling(
    bootstrap_options = c(
      "striped",
      "hover",
      "condensed"
    ),
    full_width = FALSE
  )

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

The per capita income provides a measure of the average income earned per person, reflecting the economic well-being of residents. It serves as a key indicator for assessing local economic development, standard of living, and income distribution within the community.

**Source:** ACS 2024 5-YR Estimates (table B08528)

[Download Data](data_download/PerCapitaIncome.csv){download="PerCapitaIncome.csv"}
:::