---
title: "Workforce"
editor: visual
format:
  html:
    theme: cosmo
---

## 

Please [contact us](mailto:regional@upjohn.org) if you have any questions.

## 


```{r}
#| label: EPOP
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

labor <- readRDS("data/EPOP_LFPR_UR.rds")

all_counties <- unique(labor$county)

init <- all_counties[1]

df_init <- labor %>%
  filter(county == init)

p <- plot_ly()

# County epop 
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~EPOP,
    name = init,
    line = list(width = 3, color = "#AF2C32"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

# Michigan unemployment rate
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~MI_employment_population_ratio,
    name = "Michigan",
    line = list(width = 3, dash = "dash", color = "#2B9CB4"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

buttons <- lapply(seq_along(all_counties), function(i) {

  df <- labor %>%
    filter(county == all_counties[i])

   list(
    method = "update",

    args = list(

      # Update traces
      list(
        x = list(
          df$period_date,
          df$period_date
        ),

        y = list(
          df$EPOP,
          df$MI_employment_population_ratio
        ),

        name = list(
          all_counties[i],
          "Michigan"
        ),

        customdata = list(
          df$period_txt,
          df$period_txt
        )
      ),

      # Update layout
      list(
        title = paste(
          "Employment to Population Ratio:",
          all_counties[i],
          "vs Michigan"
        )
      )
    ),

    label = all_counties[i]
  )
})

p %>%
  layout(

    title = paste(
      "Employment to Population Ratio:",
      init,
      "vs Michigan"
    ),

    xaxis = list(
      title = "Period"),

    yaxis = list(title = "Percent"),

    updatemenus = list(
      list(
        buttons = buttons,
        x = 1.25,
        y = 0.75,
        xanchor = "right"
      )
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

The employment-to-population ratio is a crucial economic indicator as it reflects the percentage of the working-age population who are employed. It provides insights into labor market health, income levels, productivity, and social stability. Policymakers and investors rely on this ratio to gauge economic well-being and make informed decisions.

**Source:** BLS LAUS, Upjohn Regional • (p) Preliminary, (e) Upjohn Estimate

[Download Data](data_download/EPOP_LFPR_UR.csv){download="EPOP_LFPR_UR.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: LFPR
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

labor <- readRDS("data/EPOP_LFPR_UR.rds")

all_counties <- unique(labor$county)

init <- all_counties[1]

df_init <- labor %>%
  filter(county == init)

p <- plot_ly()

# County epop 
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~LFPR,
    name = init,
    line = list(width = 3, color = "#AF2C32"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

# Michigan unemployment rate
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~MI_labor_force_participation_rate,
    name = "Michigan",
    line = list(width = 3, dash = "dash", color = "#2B9CB4"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

buttons <- lapply(seq_along(all_counties), function(i) {

  df <- labor %>%
    filter(county == all_counties[i])

   list(
    method = "update",

    args = list(

      # Update traces
      list(
        x = list(
          df$period_date,
          df$period_date
        ),

        y = list(
          df$LFPR,
          df$MI_labor_force_participation_rate
        ),

        name = list(
          all_counties[i],
          "Michigan"
        ),

        customdata = list(
          df$period_txt,
          df$period_txt
        )
      ),

      # Update layout
      list(
        title = paste(
          "Labor Force Participation Rate:",
          all_counties[i],
          "vs Michigan"
        )
      )
    ),

    label = all_counties[i]
  )
})

p %>%
  layout(

    title = paste(
      "Labor Force Participation Rate:",
      init,
      "vs Michigan"
    ),

    xaxis = list(
      title = "Period"),

    yaxis = list(title = "Percent"),

    updatemenus = list(
      list(
        buttons = buttons,
        x = 1.25,
        y = 0.75,
        xanchor = "right"
      )
    ),

    margin = list(
      t = 30
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

The labor force participation rate is a measure that represents the percentage of the working-age population (individuals aged 16 and older) who are either employed or actively seeking employment. Tracking the labor force participation rate provides insights on the working-age population and their engagement in the labor market.

**Source:** BLS LAUS, Upjohn Regional • (p) Preliminary, (e) Upjohn Estimate

[Download Data](data_download/EPOP_LFPR_UR.csv){download="EPOP_LFPR_UR.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: UR
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

labor <- readRDS("data/EPOP_LFPR_UR.rds")

all_counties <- unique(labor$county)

init <- all_counties[1]

df_init <- labor %>%
  filter(county == init)

p <- plot_ly()

# County unemployment rate
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~unemployment_rate,
    name = init,
    line = list(width = 3, color = "#AF2C32"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

# Michigan unemployment rate
p <- p %>%
  add_lines(
    data = df_init,
    x = ~period_date,
    y = ~MI_unemployment_rate,
    name = "Michigan",
    line = list(width = 3, dash = "dash", color = "#2B9CB4"),
    customdata = ~period_txt,
    hovertemplate = paste(
      "%{x}<br>",
      "Value: %{y:.2f}%<br>",
      "Period: %{customdata}<extra></extra>"
    )
  )

buttons <- lapply(seq_along(all_counties), function(i) {

  df <- labor %>%
    filter(county == all_counties[i])

   list(
    method = "update",

    args = list(

      # Update traces
      list(
        x = list(
          df$period_date,
          df$period_date
        ),

        y = list(
          df$unemployment_rate,
          df$MI_unemploymen_rate
        ),

        name = list(
          all_counties[i],
          "Michigan"
        ),

        customdata = list(
          df$period_txt,
          df$period_txt
        )
      ),

      # Update layout
      list(
        title = paste(
          "Unemployment Rate:",
          all_counties[i],
          "vs Michigan"
        )
      )
    ),

    label = all_counties[i]
  )
})

p %>%
  layout(

    title = paste(
      "Unemployment Rate:",
      init,
      "vs Michigan"
    ),

    xaxis = list(
      title = "Period"),

    yaxis = list(title = "Percent"),

    updatemenus = list(
      list(
        buttons = buttons,
        x = 1.25,
        y = 0.75,
        xanchor = "right"
      )
    ),

    margin = list(
      t = 30
    )
  ) %>%
  config(displayModeBar = FALSE)
```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

The unemployment rate is a measure that represents the percentage of individuals who are actively seeking employment but are currently without a job. The unemployment rate offers understanding into the availability of jobs and the overall state of the labor market within an economy.

**Source:** BLS LAUS, Upjohn Regional • (p) Preliminary, (e) Upjohn Estimate

[Download Data](data_download/EPOP_LFPR_UR.csv){download="EPOP_LFPR_UR.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: AWW
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)

qcew_data <- readRDS("data/qcew.rds")

# Ensure proper ordering
qcew_data <- qcew_data %>%
  arrange(year, quarter)

# Get counties
counties <- unique(qcew_data$county)

# Default county
default_county <- "Kalamazoo"

# Initial dataset
initial_data <- qcew_data %>%
  filter(county == default_county)

# Create plot
fig <- plot_ly()

# County bars
fig <- fig %>%
  add_trace(
    data = initial_data,
    x = ~date,
    y = ~avg_wkly_wage,
    type = "bar",
    name = default_county,
    marker = list(color = "#AF2C32"),
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "Average Weekly Wage: $%{y:,.0f}",
      "<extra></extra>"
    )
  )

# State line
fig <- fig %>%
  add_trace(
    data = initial_data,
    x = ~date,
    y = ~state_avg_wkly_wage,
    type = "scatter",
    mode = "lines",
    name = "Michigan",
    line = list(color = "#2B9CB4", dash = "dash"),
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "Michigan: $%{y:,.0f}",
      "<extra></extra>"
    )
  )

# Dropdown buttons
buttons <- lapply(seq_along(counties), function(i) {

  county_name <- counties[i]

  county_data <- qcew_data %>%
    filter(county == county_name)

  list(
    method = "update",
    args = list(
      list(
        y = list(
          county_data$avg_wkly_wage,
          county_data$state_avg_wkly_wage
        ),
        x = list(
          county_data$date,
          county_data$date
        )
      ),
      list(
        title = paste(
          "Average Weekly Wage:",
          county_name,
          "County vs Michigan"
        )
      )
    ),
    label = county_name
  )
})

# Layout
fig <- fig %>%
  layout(
    title = paste(
      "Average Weekly Wage:",
      default_county,
      "County vs Michigan"
    ),
    
    updatemenus = list(
      list(
        active = which(counties == default_county) - 1,
        buttons = buttons,
        x = 1.25,
        y = .75,
        xanchor = "right"
      )
    ),
    
    xaxis = list(title = "Date"),
    
    yaxis = list(
      title = "Average Weekly Wage ($)"
    ),
    
    barmode = "group"
  )

config(fig,
       displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Tracking the average weekly wage gives insight to the income levels and purchasing power of workers within an economy.

**Source:** BLS QCEW, Upjohn Regional • Not adjusted for inflation • (p) Preliminary

[Download Data](data_download/qcew.csv){download="qcew.csv"}
:::

::: {.plot-space}
:::

```{r}
#| label: Class of Worker
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)
library(tidyr)

# Read data
workers <- readRDS("data/classOfWorker.rds")

legend_annotations <- list(
  list(x = 1.02, y = 1.00, xref = "paper", yref = "paper",
       text = "<b>Educational Attainment</b>",
       showarrow = FALSE, xanchor = "left"),

  list(x = 1.02, y = 0.92, xref = "paper", yref = "paper",
       text = "■ Private Wage and Salary Workers",
       font = list(color = "#004A5E"),
       showarrow = FALSE, xanchor = "left"),

  list(x = 1.02, y = 0.84, xref = "paper", yref = "paper",
       text = "■ Government Workers",
       font = list(color = "#2B9CB4"),
       showarrow = FALSE, xanchor = "left"),
  
    list(x = 1.02, y = 0.76, xref = "paper", yref = "paper",
       text = "■ Self-employed in own, not incorporated, <br> business workers",
       font = list(color = "#AF2C32"),
       showarrow = FALSE, xanchor = "left"),
  
      list(x = 1.02, y = 0.76, xref = "paper", yref = "paper",
       text = "■ Unpaid family workers",
       font = list(color = "#414042"),
       showarrow = FALSE, xanchor = "left")
)


# Convert to long format
workers_long <- workers %>%
  pivot_longer(
    cols = c(
      "Private Wage and Salary Workers",
      "Government Workers",
      "Self-employed in own, not incorporated, business workers",
      "Unpaid family workers"
    ),
    names_to = "WorkerClass",
    values_to = "Percent"
  )


# Worker colors
worker_colors <- c(
  "Private Wage and Salary Workers" = "#004A5E",
  "Government Workers" = "#2B9CB4",
  "Self-employed in own, not incorporated, business workers" = "#AF2C32",
  "Unpaid family workers" = "#414042"
)

worker_labels <- c(
  "Private Wage and Salary Workers" =
    "Private Wage and Salary Workers",

  "Government Workers" =
    "Government Workers",

  "Self-employed in own, not incorporated, business workers" =
    "Self-employed in own, not incorporated,<br>business workers",

  "Unpaid family workers" =
    "Unpaid family workers"
)


# Counties
all_counties <- unique(workers_long$NAME)

init <- all_counties[1]


# Create legend annotations
legend_annotations <- list(
  list(
    x = 1.02, y = 1.00,
    xref = "paper",
    yref = "paper",
    text = "<b>Class of Worker</b>",
    showarrow = FALSE,
    xanchor = "left"
  )
)


# Add worker categories to legend annotations
y_pos <- c(.92, .84, .76, .62)

for (i in seq_along(worker_colors)) {

  worker_name <- names(worker_colors)[i]

  legend_annotations <- append(
    legend_annotations,
    list(
      list(
        x = 1.02,
        y = y_pos[i],
        xref = "paper",
        yref = "paper",
        text = paste0(
          "■ ",
          worker_labels[worker_name]
        ),
        font = list(
          color = worker_colors[i]
        ),
        showarrow = FALSE,
        xanchor = "left",
        align = "left"
      )
    )
  )
}


# Create traces
p <- plot_ly()

for (county in all_counties) {

  df_county <- workers_long %>%
    filter(NAME == county)

  for (worker in names(worker_colors)) {

    value <- df_county %>%
      filter(WorkerClass == worker) %>%
      pull(Percent)

    p <- p %>%
      add_trace(
        x = county,
        y = value,
        type = "bar",
        name = worker,
        legendgroup = worker,
        marker = list(
          color = worker_colors[worker]
        ),
        hovertemplate = paste0(
          "County: ", county,
          "<br>Share: %{y:.2f}%",
          "<extra></extra>"
        ),
        visible = (county == all_counties[1]),
        showlegend = FALSE
      )
  }
}


# Dropdown buttons
buttons <- lapply(seq_along(all_counties), function(i){

  vis <- rep(FALSE, length(all_counties) * length(worker_colors))

  start <- (i - 1) * length(worker_colors) + 1
  end <- start + length(worker_colors) - 1

  vis[start:end] <- TRUE


  list(
    method = "update",
    args = list(
      list(
        visible = vis
      ),
      list(
        title = paste(
          "Class of Worker -",
          all_counties[i]
        )
      )
    ),
    label = all_counties[i]
  )
})


# Plot
p %>%
  layout(

    title = paste(
      "Class of Worker -",
      all_counties[1]
    ),

    barmode = "stack",

    xaxis = list(
      title = ""
    ),
    
    

    yaxis = list(
      title = "Percent",
      range = c(0, 100),
      showgrid = TRUE,
      gridcolor = "lightgray",
      gridwidth = 1
    ),


    updatemenus = list(
      list(
        buttons = buttons,
      active = 0,
      x = 1.05,
      xanchor = "left",
      y = 0.5,
      yanchor = "top"
      )
    ),


    showlegend = FALSE,
    annotations = legend_annotations <- c(
      legend_annotations
    ),


    margin = list(
      r = 300
    )

  ) %>%
  config(displayModeBar = FALSE)


```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Understanding the class of worker in a county is crucial for analyzing labor market dynamics and economic structure. It offers insights into employment patterns, including self-employment, wage and salary jobs, and their contributions to economic activity and workforce composition.

**Source:** ACS 2024 5-YR Estimates (table B08528)

[Download Data](data_download/averageAnnualWage.csv){download="averageAnnualWage.csv"}
:::

##

::: plot-space
:::


```{r}
#| label: Educational Attainment
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)
library(tidyr)

# ---------------------------------------------------------
# Read data
# ---------------------------------------------------------

S1501_wide <- readRDS("data/eduAttainment.rds")


# ---------------------------------------------------------
# Southwest Michigan county order
# ---------------------------------------------------------

county_order <- c(
  "Kalamazoo",
  "Berrien",
  "Branch",
  "Calhoun",
  "Cass",
  "St. Joseph",
  "Van Buren"
)


# ---------------------------------------------------------
# Education category order
# ---------------------------------------------------------

edu_levels <- c(
  "Less than 9th grade",
  "9th to 12th grade, no diploma",
  "High school graduate (includes equivalency)",
  "Some college, no degree",
  "Associate's degree",
  "Bachelor's degree",
  "Graduate or professional degree"
)


# ---------------------------------------------------------
# Convert to long format
# ---------------------------------------------------------

S1501_long <- S1501_wide %>%
  pivot_longer(
    cols = all_of(edu_levels),
    names_to = "Education",
    values_to = "Percent"
  ) %>%
  mutate(
    Education = factor(
      Education,
      levels = edu_levels
    )
  )


# ---------------------------------------------------------
# Sequential education colors
# ---------------------------------------------------------

edu_colors <- c(
  "Less than 9th grade" = "#cf203e",
  "9th to 12th grade, no diploma" = "#ba2a4e",
  "High school graduate (includes equivalency)" = "#a5345f",
  "Some college, no degree" = "#903f6f",
  "Associate's degree" = "#7a497f",
  "Bachelor's degree" = "#655390",
  "Graduate or professional degree" = "#505da0"
)


# ---------------------------------------------------------
# Legend labels
# ---------------------------------------------------------

legend_labels <- c(
  "Less than 9th grade",
  "9th to 12th grade, no diploma",
  "High school graduate",
  "Some college, no degree",
  "Associate's degree",
  "Bachelor's degree",
  "Graduate or professional degree"
)


# ---------------------------------------------------------
# Counties
# Southwest Michigan first, then remaining counties
# ---------------------------------------------------------

all_counties <- unique(
  as.character(S1501_long$NAME)
)

all_counties <- c(
  county_order[county_order %in% all_counties],
  sort(setdiff(all_counties, county_order))
)


# ---------------------------------------------------------
# Initial county
# ---------------------------------------------------------

init <- all_counties[1]

df_init <- S1501_long %>%
  filter(NAME == init) %>%
  arrange(Education)


# ---------------------------------------------------------
# Create legend annotations
# ---------------------------------------------------------

legend_annotations <- list(
  list(
    x = 1.02,
    y = 1.00,
    xref = "paper",
    yref = "paper",
    text = "<b>Educational Attainment</b>",
    showarrow = FALSE,
    xanchor = "left"
  )
)

y_pos <- c(
  0.92,
  0.84,
  0.76,
  0.68,
  0.60,
  0.52,
  0.44
)

for (i in seq_along(edu_levels)) {

  legend_annotations <- append(
    legend_annotations,
    list(
      list(
        x = 1.02,
        y = y_pos[i],
        xref = "paper",
        yref = "paper",

        text = paste0(
          "■ ",
          legend_labels[i]
        ),

        font = list(
          color = edu_colors[edu_levels[i]]
        ),

        showarrow = FALSE,
        xanchor = "left"
      )
    )
  )
}


# ---------------------------------------------------------
# Create column chart
# ---------------------------------------------------------

p <- plot_ly(
  data = df_init,
  x = ~Education,
  y = ~Percent,
  type = "bar",

  marker = list(
    color = edu_colors[
      as.character(df_init$Education)
    ]
  ),

  hovertemplate = paste0(
    "County: ", init,
    "<br>Educational Attainment: %{x}",
    "<br>Share: %{y:.2f}%",
    "<extra></extra>"
  )
)


# ---------------------------------------------------------
# County dropdown
# ---------------------------------------------------------

buttons <- lapply(
  seq_along(all_counties),
  function(i) {

    df <- S1501_long %>%
      filter(NAME == all_counties[i]) %>%
      arrange(Education)

    list(
      method = "update",

      args = list(

        # Update bar data
        list(
          x = list(df$Education),
          y = list(df$Percent),

          marker = list(
            color = edu_colors[
              as.character(df$Education)
            ]
          ),

          hovertemplate = paste0(
            "County: ", all_counties[i],
            "<br>Educational Attainment: %{x}",
            "<br>Share: %{y:.2f}%",
            "<extra></extra>"
          )
        ),

        # Update title
        list(
          title = paste(
            "Educational Attainment -",
            all_counties[i]
          )
        )
      ),

      label = all_counties[i]
    )
  }
)


# ---------------------------------------------------------
# Render chart
# ---------------------------------------------------------

p %>%
  layout(

    title = paste(
      "Educational Attainment -",
      init
    ),

    # Regular column chart
    barmode = "group",

    # X-axis
    xaxis = list(
      title = "",
      showticklabels = FALSE,

      categoryorder = "array",
      categoryarray = edu_levels
    ),

    # Y-axis
    yaxis = list(
      title = "Percent",
      ticksuffix = "%",
      showgrid = TRUE,
      gridcolor = "lightgray",
      gridwidth = 1
    ),

    # Hide Plotly's default legend
    showlegend = FALSE,

    # Custom annotation legend
    annotations = legend_annotations,

    # County dropdown
    updatemenus = list(
      list(
        type = "dropdown",
        buttons = buttons,
        active = 0,
        x = 1.25,
        y = 0.35,
        xanchor = "left"
      )
    ),

    # Space for legend and dropdown
    margin = list(
      r = 280,
      b = 100
    )

  ) %>%
  config(
    displayModeBar = FALSE
  )
```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Educational attainment strongly correlates with income levels, workforce productivity, and economic growth. It helps gauge the skill level and potential earning capacity of a population.

**Source:** ACS 2024 5-YR Estimates (table S1501) • Population 25 and older

[Download Data](data_download/eduAttainment.csv){download="eduAttainment.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Income by Sex and Edcational Attainment
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(tidyr)
library(plotly)

income <- readRDS("data/incomeBySexAndEducationalAttainment.rds")

income$variable <- factor(
  income$variable,
  levels = c(
    "Less than high school graduate",
    "High school graduate (includes equivalency)",
    "Some college or associate's degree",
    "Bachelor's degree",
    "Graduate or professional degree"
  )
)

income <- income %>%
  mutate(
    Display = recode(
      variable,
      "Less than high school graduate" = "Less Than H.S.",
      "High school graduate (includes equivalency)" = "H.S.",
      "Some college or associate's degree" = "Some College/Associates",
      "Bachelor's degree" = "Bachelor's",
      "Graduate or professional degree" = "Graduate/Professional"
    )
  )

income$Display <- factor(
  income$Display,
  levels = c(
    "Less Than H.S.",
    "H.S.",
    "Some College/Associates",
    "Bachelor's",
    "Graduate/Professional"
  )
)

income_colors <- c(
  "Male" = "#004A5E",
  "Female" = "#2B9CB4"
)

counties <- unique(income$NAME)

init <- counties[1]

df_init <- income %>%
  filter(NAME == init)

p <- plot_ly()

p <- p %>%
  add_bars(
    data = df_init,
    x = ~Display,
    y = ~`Male Income`,
    name = "Male",
    marker = list(
      color = income_colors["Male"]
    )
  ) %>%
  add_bars(
    data = df_init,
    x = ~Display,
    y = ~`Female Income`,
    name = "Female",
    marker = list(
      color = income_colors["Female"]
    )
  )

buttons <- lapply(seq_along(counties), function(i) {

  df <- income %>%
    filter(NAME == counties[i])

  list(
    method = "update",
    args = list(
      list(
        x = list(df$Display, df$Display),
        y = list(
          df$`Male Income`,
          df$`Female Income`
        )
      ),
      list(
        title = paste(
          "Income by Educational Attainment:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )
})

p %>%
  layout(
    barmode = "group",

    title = paste(
      "Income by Educational Attainment:",
      init
    ),

    xaxis = list(
      title = ""
    ),

    yaxis = list(
      title = "Income",
      tickprefix = "$"
    ),

    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.25,
        y = 0.75,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 250
    )
  ) %>%
  config(displayModeBar = FALSE)
```

::: {.callout-note collapse="true" appearance="minimal"}
## Description, Source, and Data Download.

Understanding income disparities by educational attainment and sex provides insights into the effectiveness of educational systems in equipping individuals with skills for higher-paying jobs and highlights potential gender-based income gaps, which can be indicative of gender equality issues.

**Source:** ACS 2024 5-YR Estimates (table B20004) • Population 25 and older with earnings

[Download Data](data_download/incomeBySexAndEducationalAttainment.csv){download="incomeBySexAndEducationalAttainment.csv"}
:::
