---
title: "Housing and Transportation"
editor: visual
format:
  html:
    theme: cosmo
---

## 

Please [contact us](mailto:regional@upjohn.org) if you have any questions.

## 

::: plot-space
:::

```{r}
#| label: Travel Time to Work
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

travel <- readRDS("data/travelTimeToWork.rds")

travel$TimeLabel <- factor(
  travel$TimeLabel,
  levels = c(
    "less than 10 minutes",
    "10 to 29 minutes",
    "30 to 59 minutes",
    "60 or more minutes"
  )
)

counties <- unique(travel$NAME)

init <- counties[1]

df_init <- travel %>%
  filter(NAME == init)

travel_colors <- c(
  "less than 10 minutes" = "#FFE293",
  "10 to 29 minutes" = "#FDBA4D",
  "30 to 59 minutes" = "#F68C3D",
  "60 or more minutes" = "#AF2C32"
)

p <- plot_ly(
  data = df_init,
  x = ~TimeProportion,
  y = ~TimeLabel,
  type = "bar",
  orientation = "h",
  hovertemplate = "%{y}<br>%{x:.1f}%<extra></extra>",
  marker = list(
    color = travel_colors[df_init$TimeLabel]
  )
)

buttons <- lapply(seq_along(counties), function(i) {

  df <- travel %>%
    filter(NAME == counties[i])

  list(
    method = "update",
    args = list(
      list(
        x = list(df$TimeProportion),
        y = list(df$TimeLabel),
        marker = list(
          color = travel_colors[df$TimeLabel]
        )
      ),
      list(
        title = paste(
          "Travel Time to Work:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )
})

p %>%
  layout(
    title = paste(
      "Travel Time to Work:",
      init
    ),

    xaxis = list(
      title = "Percent of Workers 16 Years and Older"
    ),

    yaxis = list(
      title = ""
    ),

    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.15,
        y = 0.85,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 150
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

Travel time to work is significant as it directly impacts worker productivity and transportation costs. It also informs urban planning and labor market dynamics, influencing economic development and accessibility to jobs.

**Source:** ACS 2024 5-YR Estimates (table B08135) • Travel time one way

[Download Data](data_download/travelTimeToWork.csv){download="travelTimeToWork.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Means of Transportation to Work
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

transport <- readRDS("data/meansOfTransportationToWork.rds")

transport$Category <- factor(
  transport$Category,
  levels = c(
    "Car, truck, or van - drove alone",
    "Car, truck, or van - carpooled",
    "Public transportation (excluding taxicab)",
    "Walked",
    "Taxicab, motorcycle, bicycle, or other means",
    "Worked from home"
  )
)

transport <- transport %>%
  mutate(
    Display = recode(
      Category,
      "Car, truck, or van - drove alone" = "Drive Alone",
      "Car, truck, or van - carpooled" = "Carpool",
      "Public transportation (excluding taxicab)" = "Public Transit",
      "Walked" = "Walk",
      "Taxicab, motorcycle, bicycle, or other means" = "Other",
      "Worked from home" = "Work From Home"
    )
  )

transport$Display <- factor(
  transport$Display,
  levels = c(
    "Drive Alone",
    "Carpool",
    "Public Transit",
    "Walk",
    "Other",
    "Work From Home"
  )
)

transport_colors <- c(
  "Drive Alone" = "#FFE293",
  "Carpool" = "#FDBA4D",
  "Public Transit" = "#F68C3D",
  "Walk" = "#F15A29",
  "Other" = "#AF2C32",
  "Work From Home" = "#822B1A"
)

counties <- unique(transport$NAME)

init <- counties[1]

df_init <- transport %>%
  filter(NAME == init)

p <- plot_ly(
  data = df_init,
  x = ~Proportion,
  y = ~Display,
  type = "bar",
  orientation = "h",
  marker = list(
    color = transport_colors[df_init$Display]
  ),
  hovertemplate = "%{y}<br>%{x:.1f}%<extra></extra>"
)

buttons <- lapply(seq_along(counties), function(i) {

  df <- transport %>%
    filter(NAME == counties[i])

  list(
    method = "update",
    args = list(
      list(
        x = list(df$Proportion),
        y = list(df$Display),
        marker = list(
          color = transport_colors[df$Display]
        )
      ),
      list(
        title = paste(
          "Means of Transportation to Work:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )
})

p %>%
  layout(
    title = paste(
      "Means of Transportation to Work:",
      init
    ),

    xaxis = list(
      title = "Percent of Workers 16 Years and Older"
    ),

    yaxis = list(
      title = "",
      categoryorder = "array",
      categoryarray = rev(levels(transport$Display))
    ),


    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.15,
        y = 0.85,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 150
    )
  ) %>%
  config(displayModeBar = FALSE)



```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

Understanding how people commute to work helps assess the efficiency of transportation infrastructure, which can impact productivity and business operations. It also influences urban planning and investment decisions, as well as energy consumption and environmental considerations. Additionally, tracking transportation methods can reveal trends in workforce mobility and inform policies that affect job accessibility and economic opportunities.

**Source:** ACS 2024 5-YR Estimates (table B08128)

[Download Data](data_download/meansOfTransportationToWork.csv){download="meansOfTransportationToWork.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Age of Housing Stock
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

housing <- readRDS("data/ageOfHousingStock.rds")

housing$variable <- factor(
  housing$variable,
  levels = c(
    "2020 or later",
    "2010 to 2019",
    "2000 to 2009",
    "1980 to 1999",
    "1960 to 1979",
    "1940 to 1959",
    "1939 or earlier"
  )
)


housing_colors <- c(
  "2020 or later" = "#FFE293",
  "2010 to 2019" = "#FDBA4D",
  "2000 to 2009" = "#F68C3D",
  "1980 to 1999" = "#F15A29",
  "1960 to 1979" = "#DC4133",
  "1940 to 1959" = "#AF2C32",
  "1939 or earlier" = "#822B1A"
)

counties <- unique(housing$NAME)

init <- counties[1]

df_init <- housing %>%
  filter(NAME == init)

p <- plot_ly(
  data = df_init,
  x = ~estimate,
  y = ~variable,
  type = "bar",
  orientation = "h",
  marker = list(
  color = housing_colors[as.character(df_init$variable)]
  ),
  hovertemplate = "%{y}<br>%{x:.1f}%<extra></extra>"
)

buttons <- lapply(seq_along(counties), function(i) {

  df <- housing %>%
    filter(NAME == counties[i])

  list(
    method = "update",
    args = list(
      list(
        x = list(df$estimate)
      ),
      list(
        title = paste(
          "Age of Housing Stock:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )
})


p %>%
  layout(
    title = paste(
      "Age of Housing Stock:",
      init
    ),

    xaxis = list(
      title = "Percent of Total Housing Stock"
    ),

    yaxis = list(
      title = "",
      categoryorder = "array",
      categoryarray = rev(levels(housing$variable))
    ),

    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.15,
        y = 0.85,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 150
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

Understanding the age of the housing stock reveals the state of infrastructure and housing quality. Older housing may require more maintenance and renovations, which can drive construction and home improvement industries, thereby boosting economic activity. Additionally, it can inform urban development and land use policies, impacting property values and real estate investments. Lastly, the age of the housing stock can provide insights into population trends, migration patterns, and housing affordability, all of which have significant economic implications.

**Source:** ACS 2024 5-YR Estimates (table S2504)

[Download Data](data_download/ageOfHousingStock.csv){download="ageOfHousingStock.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Internet Access by Household
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(plotly)

internet <- readRDS("data/internetAccess.rds")

internet <- internet %>%
  mutate(
    Category = recode(
      variable,
      "Dial-up with no other type of Internet subscription" = "Dial-Up",
      "Broadband of any type" = "Broadband",
      "Without an Internet subscription" = "No Internet"
    )
  )

internet_colors <- c(
  "Broadband" = "#2B9CB4",
  "Dial-Up" = "#AF2C32",
  "No Internet" = "#414042"
)

counties <- unique(internet$NAME)

init <- counties[1]

df_init <- internet %>%
  filter(NAME == init)


p <- plot_ly(
  data = df_init,
  labels = ~Category,
  values = ~estimate,
  type = "pie",
  sort = FALSE,
  direction = "clockwise",
  marker = list(
    colors = internet_colors[df_init$Category]
  ),
  textinfo = "none",
  hovertemplate = "<b>%{label}</b><br>%{value:.1f}%<extra></extra>"
)


buttons <- lapply(seq_along(counties), function(i) {

  df <- internet %>%
    filter(NAME == counties[i])

  list(
    method = "update",
    args = list(
      list(
        labels = list(df$Category),
        values = list(df$estimate),
        marker = list(
          colors = internet_colors[df$Category]
        )
      ),
      list(
        title = paste(
          "Internet Access:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )
})


p %>%
  layout(
    title = paste(
      "Internet Access:",
      init
    ),

    showlegend = TRUE,

    legend = list(
      x = 1.00,
      y = 1.00,
      xanchor = "left"
    ),

    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.20,
        y = 0.85,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 250
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

Knowledge of people's internet access is an important economic indicator because it reflects digital connectivity, which is increasingly vital for economic participation. It influences business competitiveness and innovation, as companies depend on digital platforms for operations and sales. Internet access also affects workforce capabilities, as remote work and online education become more prevalent. Finally, understanding how people access the internet plays a role in bridging economic disparities, as access can be a factor in determining opportunities and socioeconomic outcomes.

**Source:** ACS 2024 5-YR Estimates (table S2801) • Totals may not sum to 100% due to rounding.

[Download Data](data_download/internetAccess.csv){download="internetAccess.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Building Permits
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)
library(tidyr)

permits <- readRDS(
  "data/buildingPermits.rds"
)

# Southwest Michigan counties
county_order <- c(
  "Kalamazoo",
  "Berrien",
  "Branch",
  "Calhoun",
  "Cass",
  "St. Joseph",
  "Van Buren"
)

# Convert building permit categories to long format
permits_long <- permits %>%
  pivot_longer(
    cols = c(
      `1-unit`,
      `2-units`,
      `3-4 units`,
      `5+ units`
    ),
    names_to = "Building Type",
    values_to = "Permits"
  ) %>%
  mutate(
    Year = as.numeric(Survey),
    Permits = as.numeric(Permits)
  )

# Building type order
building_order <- c(
  "1-unit",
  "2-units",
  "3-4 units",
  "5+ units"
)

permits_long$`Building Type` <- factor(
  permits_long$`Building Type`,
  levels = building_order
)

# Get all counties
counties <- unique(
  as.character(permits_long$County)
)

# Southwest Michigan counties first,
# followed by all other counties alphabetically
counties <- c(
  county_order[county_order %in% counties],
  sort(setdiff(counties, county_order))
)

# Initial county
init <- counties[1]

# Initial data
df_init <- permits_long %>%
  filter(County == init)

# Create chart
p <- plot_ly(
  data = df_init,
  x = ~Year,
  y = ~Permits,
  color = ~`Building Type`,
  colors = c(
    "#004A5E",
    "#2B9CB4",
    "#AF2C32",
    "#414042"
  ),
  type = "scatter",
  mode = "lines+markers"
)

# County dropdown
buttons <- lapply(counties, function(county) {

  df <- permits_long %>%
    filter(County == county) %>%
    arrange(Year, `Building Type`)

  list(
    method = "update",

    args = list(

      list(
        x = list(
          df$Year[df$`Building Type` == "1-unit"],
          df$Year[df$`Building Type` == "2-units"],
          df$Year[df$`Building Type` == "3-4 units"],
          df$Year[df$`Building Type` == "5+ units"]
        ),

        y = list(
          df$Permits[df$`Building Type` == "1-unit"],
          df$Permits[df$`Building Type` == "2-units"],
          df$Permits[df$`Building Type` == "3-4 units"],
          df$Permits[df$`Building Type` == "5+ units"]
        )
      ),

      list(
        title = paste(
          "Building Permits:",
          county
        )
      )
    ),

    label = county
  )
})

# Layout
p %>%
  layout(

    title = paste(
      "Building Permits:",
      init
    ),

    xaxis = list(
      title = "Year",
      dtick = 1
    ),

    yaxis = list(
      title = "Housing Units",
      separatethousands = TRUE
    ),

    legend = list(
      orientation = "v",
      x = 1.02,
      y = 0.95,
      xanchor = "left",
      yanchor = "top"
    ),

    updatemenus = list(
      list(
        type = "dropdown",
        buttons = buttons,
        active = 0,
        x = 1.15,
        y = 0.65,
        xanchor = "left"
      )
    ),

    margin = list(
      r = 220
    )
  ) %>%
  config(
    displayModeBar = FALSE
  )

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

The number of building permits issued can be crucial for gauging the vitality of the construction sector, offering insights into real estate investment trends and potential economic growth. Fluctuations in permit numbers can signal changes in job creation within construction and related industries, serving as a barometer for broader economic activity. A decline in permits may indicate economic challenges, affecting employment and the overall economic environment.

**Source:** Residential Building Permits Survey County ASC II Files.

[Download Data](data_download/buildingPermits.csv){download="buildingPermits.csv"}
:::

##

::: plot-space
:::

```{r}
#| label: Monthly Income Needed to Afford Housing
#| echo: false
#| message: false
#| warning: false

library(plotly)
library(dplyr)
library(tidyr)

housing <- readRDS("data/monthlyIncomeNeededToAffordHousing.rds")

housing_long <- housing %>%
  pivot_longer(
    cols = c(`Median Cost`, `Income Needed to Afford Cost`),
    names_to = "Measure",
    values_to = "Value"
  )

housing_long$Measure <- factor(
  housing_long$Measure,
  levels = c(
    "Median Cost",
    "Income Needed to Afford Cost"
  )
)

colors <- c(
  "Median Cost" = "#004A5E",
  "Income Needed to Afford Cost" = "#F68C3D"
)

counties <- unique(housing_long$NAME)

init <- counties[1]

df_init <- housing_long %>%
  filter(NAME == init)

p <- plot_ly()

p <- p %>%
  add_bars(
    data = df_init %>% filter(Measure == "Median Cost"),
    x = ~variable,
    y = ~Value,
    name = "Median Cost",
    marker = list(color = colors["Median Cost"])
  ) %>%
  add_bars(
    data = df_init %>% filter(Measure == "Income Needed to Afford Cost"),
    x = ~variable,
    y = ~Value,
    name = "Income Needed to Afford Cost",
    marker = list(color = colors["Income Needed to Afford Cost"])
  )

buttons <- lapply(seq_along(counties), function(i) {

  df <- housing_long %>%
    filter(NAME == counties[i])

  median_df <- df %>%
    filter(Measure == "Median Cost")

  income_df <- df %>%
    filter(Measure == "Income Needed to Afford Cost")

  list(
    method = "update",
    args = list(
      list(
        x = list(median_df$variable,
                 income_df$variable),
        y = list(median_df$Value,
                 income_df$Value)
      ),
      list(
        title = paste(
          "Housing Costs:",
          counties[i]
        )
      )
    ),
    label = counties[i]
  )

})

p %>%
  layout(
    barmode = "group",

    title = paste(
      "Monthly Income Needed to Afford Housing:",
      init
    ),

    xaxis = list(
      title = ""
    ),

    yaxis = list(
      tickprefix = "$",
      separatethousands = TRUE
    ),

    legend = list(
      orientation = "v",
      x = 1.02,
      y = 0.95,
      xanchor = "left",
      yanchor = "top"
    ),

    updatemenus = list(
      list(
        buttons = buttons,
        active = 0,
        x = 1.15,
        y = 0.65,
        xanchor = "left"
      )
    ),

    annotations = list(
      list(
        text = "<b>Source</b> ACS 2024 5-Year Estimates (DP04), Upjohn Regional",
        xref = "paper",
        yref = "paper",
        x = 0,
        y = -0.30,
        showarrow = FALSE,
        xanchor = "left",
        font = list(size = 10)
      )
    ),

    margin = list(
      r = 180
    )
  ) %>%
  config(displayModeBar = FALSE)

```

::: {.callout-note collapse="true" appearance="minimal"}
## Definition, Data Source and Download

Knowing the median income needed to afford rent and housing costs is essential to evaluate how housing affordability impacts on individuals and the overall economy. This knowledge informs policymakers on ways to address potential disparities, guiding strategies to enhance housing accessibility and promote sustainable economic development.

**Source:** ACS 2024 5-YR Estimates (table Table DP04) • Totals may not sum to 100% due to rounding.

[Download Data](data_download/monthlyIncomeNeededToAffordHousing.csv){download="internetAccess.csv"}
:::
